### 2017-09-19 analysis/paper/saa2018_general_sessions_word2vec.Rmd
what does word2vec do?

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(tm)
library(topicmodels)
library(readxl)
library(Rmpfr)
library(tidyverse)
```

# Introduction

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

general_2018_papers$geo <- 
  unlist(purrr::map(general_2018_papers$`Geographic Focus`, ~clean_geo(.x)))

general_2018_papers$kw1 <- 
  unlist(purrr::map(general_2018_papers$Keyword1, ~clean_geo(.x)))

general_2018_papers$kw2 <- 
  unlist(purrr::map(general_2018_papers$Keyword2, ~clean_geo(.x)))

general_2018_papers$kw3 <- 
  unlist(purrr::map(general_2018_papers$Keyword3, ~clean_geo(.x)))

general_2018_papers$text_to_model <- 
  with(general_2018_papers, paste(geo, kw1, kw2, kw3))
```

```{r}
# clean text to model
general_2018_papers$text_to_model  <- 
  stringr::str_trim(tolower(gsub("\\/|:|&", " ", general_2018_papers$text_to_model)))

general_2018_papers$text_to_model  <- 
  gsub(" and | archaeology | analysis |\r\n", " ", general_2018_papers$text_to_model)

```

```{r}
# Create a corpus object from the text to mode

general_2018_papers_corpus <- tm::Corpus(tm::VectorSource(general_2018_papers$text_to_model ))

ndocs <- length(general_2018_papers_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * .5

# create a document term matrix
general_2018_papers_tm <- tm::DocumentTermMatrix(general_2018_papers_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                     wordLengths=c(4,30),  
                                                    removeNumbers = TRUE, 
                                                     bounds = list(global = c(minTermFreq, maxTermFreq)),
                                                    removePunctuation = TRUE))
```

```{r}
library(wordVectors)
library(magrittr)
library(stringr)

# Write abstracts out to text files
lapply(1:length(general_2018_papers$Abstract),
     function(i) write.table(tolower(str_trim(general_2018_papers$Abstract[i])),
                  file = paste0("abstracts/", general_2018_papers$`Abstract Id`[i], ".txt"),
                  quote = FALSE, 
                  row.names = FALSE,
                  col.names = FALSE))

length(list.files("abstracts/")) #178

prep_word2vec(origin = "abstracts/",
              destination = "abstracts/abstracts.txt",
              lowercase = TRUE,
              bundle_ngrams = 2)

model = train_word2vec("abstracts/abstracts.txt","abstracts/abstracts.bin",
                       vectors=200,
                       threads=4,
                       window=12,
                       iter=5,
                       negative_samples=0)

# model = read.vectors("cookbook_vectors.bin")

plot(model,perplexity=50)

set.seed(10)
centers = 50
clustering = kmeans(model,
                    centers = centers,
                    iter.max = 40)

sapply(sample(1:centers,10),function(n) {
  names(clustering$cluster[clustering$cluster==n][1:10])
})

```

```{r}
library(stringr)
library(text2vec)

itn <- map(general_2018_papers$Abstract, ~itoken(.x))

it = itoken(general_2018_papers$Abstract, progressbar = FALSE)
v = create_vocabulary(it) %>% prune_vocabulary(doc_proportion_max = 0.1, term_count_min = 5)
vectorizer = vocab_vectorizer(v)

# Jaccard similarity
# they will be in the same space because we use same vectorizer
# hash_vectorizer will also work fine
dtm1 = create_dtm(it1, vectorizer)

dtm2 = create_dtm(it2, vectorizer)

# Once we have representation of documents in vector space we are almost done. One thing remains - call sim2():

d1_d2_jac_sim = sim2(dtm1, dtm2, method = "jaccard", norm = "none")
# Check result:

dim(d1_d2_jac_sim)
## [1] 300 200
d1_d2_jac_sim[1:2, 1:5]
```

```{r}
library(stringr)
library(text2vec)
data("movie_review")
# select 500 rows for faster running times
movie_review = movie_review[1:500, ]
prep_fun = function(x) {
  x %>% 
    # make text lower case
    str_to_lower %>% 
    # remove non-alphanumeric symbols
    str_replace_all("[^[:alnum:]]", " ") %>% 
    # collapse multiple spaces
    str_replace_all("\\s+", " ")
}
movie_review$review_clean = prep_fun(movie_review$review)

doc_set_1 = movie_review[1:300, ]
it1 = itoken(doc_set_1$review_clean, progressbar = FALSE)

# specially take different number of docs in second set
doc_set_2 = movie_review[301:500, ]
it2 = itoken(doc_set_2$review_clean, progressbar = FALSE)

it = itoken(movie_review$review_clean, progressbar = FALSE)
v = create_vocabulary(it) %>% prune_vocabulary(doc_proportion_max = 0.1, term_count_min = 5)
vectorizer = vocab_vectorizer(v)

# Jaccard similarity
# they will be in the same space because we use same vectorizer
# hash_vectorizer will also work fine
dtm1 = create_dtm(it1, vectorizer)
dim(dtm1)
## [1]  300 2338
dtm2 = create_dtm(it2, vectorizer)
dim(dtm2)
## [1]  200 2338
# Once we have representation of documents in vector space we are almost done. One thing remains - call sim2():

d1_d2_jac_sim = sim2(dtm1, dtm2, method = "jaccard", norm = "none")
# Check result:

dim(d1_d2_jac_sim)
## [1] 300 200
d1_d2_jac_sim[1:2, 1:5]

```

```{r}
m  <- as.matrix(general_2018_papers_tm)
distMatrix <- dist(m, method="euclidean")

groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k=70)

# see what cluster each abstract is in:
cuts <- cutree(groups, k = 50)
```

```{r}
# put topics back onto abstract spreadsheet
# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
# general_2018$`Abstract Id`

general_2018_papers_with_topics <- 
  bind_cols(general_2018_papers, data_frame(cuts))

write.csv(general_2018_papers_with_topics, "../data/derived_data/general_2018_papers_with_clusters.csv")
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```