### 2017-09-20 analysis/paper/saa2018_general_sessions_posters_hclust.Rmd
thinking about scheduling

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(tm)
library(topicmodels)
library(readxl)
library(Rmpfr)
library(tidyverse)
```

# Introduction

```{r eval=TRUE}
# read in the abstracts
general_2018 <- read_excel("../data/raw_data/General Abstracts_2018.xlsx")

general_2018_posters <- general_2018 %>% filter(grepl("Poster", `Presentation Format`))
general_2018_papers  <- general_2018 %>% filter(grepl("Paper", `Presentation Format`))

# clean geo
clean_geo <- function(x){
  tmp <- stringr::str_trim(tolower(gsub("[[:punct:]]|\r\n| and ", " ", x)))
  tmp <- gsub("other", "", tmp)
  tmp <- gsub(" ", "_", tmp)
  tmp <- gsub("_{2}", "_", tmp)
  tmp
}

general_2018_posters$geo <- 
  unlist(purrr::map(general_2018_posters$`Geographic Focus`, ~clean_geo(.x)))

general_2018_posters$kw1 <- 
  unlist(purrr::map(general_2018_posters$Keyword1, ~clean_geo(.x)))

general_2018_posters$kw2 <- 
  unlist(purrr::map(general_2018_posters$Keyword2, ~clean_geo(.x)))

general_2018_posters$kw3 <- 
  unlist(purrr::map(general_2018_posters$Keyword3, ~clean_geo(.x)))

general_2018_posters$text_to_model <- 
  with(general_2018_posters, paste(geo, kw1, kw2, kw3))
```

```{r}
# clean text to model
general_2018_posters$text_to_model  <- 
  stringr::str_trim(tolower(gsub("\\/|:|&", " ", general_2018_posters$text_to_model)))

general_2018_posters$text_to_model  <- 
  gsub(" and | archaeology | analysis |\r\n", " ", general_2018_posters$text_to_model)

```

```{r}
# Create a corpus object from the text to mode

general_2018_posters_corpus <- tm::Corpus(tm::VectorSource(general_2018_posters$text_to_model ))

ndocs <- length(general_2018_posters_corpus)
# ignore extremely rare words i.e. terms that appear in less then 1% of the documents
minTermFreq <- ndocs * 0.01
# ignore overly common words i.e. terms that appear in more than 50% of the documents
maxTermFreq <- ndocs * 0.5

# create a document term matrix
general_2018_posters_tm <- tm::DocumentTermMatrix(general_2018_posters_corpus, 
                                     control = list(stemming = TRUE, 
                                                    stopwords = TRUE,
                                                    wordLengths=c(4,30),  
                                                    removeNumbers = TRUE, 
                                                    bounds = list(global = c(minTermFreq, maxTermFreq)),
                                                    removePunctuation = TRUE))
```

Distance matrix

```{r}
m  <- as.matrix(general_2018_posters_tm)
distMatrix <- dist(m, method="euclidean")
```

How many clusters?

```{r}
library(apcluster)
d.apclus <- apcluster(negDistMat(r=2), m)
k <- length(d.apclus@clusters)

windows()
heatmap(d.apclus)
```

Cluster analysis and dendrogram

```{r}
groups <- hclust(distMatrix,method="ward.D")
plot(groups, cex=0.9, hang=-1)
rect.hclust(groups, k)

# see what cluster each abstract is in:
cuts <- cutree(groups, k)
```

```{r}
library(dynamicTreeCut)
maxCoreScatter <- 0.99
minGap <- (1 - maxCoreScatter) * 0.01
dynamicCut <- cutreeDynamic(groups, 
                           minClusterSize=4, 
                           method="hybrid", 
                           distM=as.matrix(dist(m, method="euclidean")), 
                           deepSplit=4, 
                           maxCoreScatter=maxCoreScatter, 
                           maxAbsCoreScatter=NULL, 
                           minAbsGap=NULL)

hist(table(dynamicCut))

xfreqs_per_cluster <- 
dynamicCut %>%
  data_frame(dynamicCut = .,
             freq = names(.)) %>% 
  group_by(dynamicCut) %>% 
  tally(sort=TRUE)

hist(xfreqs_per_cluster$n)

library(dendextend)
hc <- groups
dend <- as.dendrogram(hc)
heights_per_k.dendrogram(dend)

unnest <- function(x) { # from Vlo's answer
  if(is.null(names(x))) x
  else c(list(all=unname(unlist(x))), do.call(c, lapply(x, unnest)))
}

hc <- groups
cuts2 <- hc$height + 1e-9
# my experiment
for(i in seq_along(cuts2)){
  n <- attr(cut(as.dendrogram(groups), h=3)$lower[[i]], "members")
  print(n)
}

my_list <- cut(as.dendrogram(groups), h=3)
map(my_list, unnest(.x)$lower)

map(cut(as.dendrogram(groups), h=3), ~attr(.x$lower, "members"))

xcuts <- cutree(groups, k)

xfreqs_per_cluster <- 
xcuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally(sort=TRUE)

xmax <- max(xfreqs_per_cluster$n)
xmin <- min(xfreqs_per_cluster$n)

max_size <- 16
min_size <- 4
n <- 0

while(xmax <= max_size &  xmin >= min_size & n <= length(xcuts)){
  n <-  n + 1
  ycuts <- cutree(groups, n)

xfreqs_per_cluster <- 
xcuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally(sort=TRUE)

xmax <- max(xfreqs_per_cluster$n)
xmin <- min(xfreqs_per_cluster$n)

  xcuts <- cutree(groups, n)

}

```

```{r}

hc <- groups

hc$size

unnest <- function(x) { # from Vlo's answer
  if(is.null(names(x))) x
  else c(list(all=unname(unlist(x))), do.call(c, lapply(x, unnest)))
}

cuts2 <- hc$height + 1e-9

biggest <- 50
smallest <- 0

max_size <- 16
min_size <- 4

i <- 0

while(smallest < min_size & 
      biggest > max_size &
      i <= length(cuts2)){
  h_i <- cuts2[i <- i+1]
  if(i > length(cuts2)){
    warning("Couldn't find a cluster big enough.")
  }
  else  smallest <- 
           Reduce(min, 
                  lapply(X = unnest(cut(as.dendrogram(hc), h=h_i)$lower), 
                         FUN = attr, which = "members") ) # from lukeA's comment

  biggest <- 
           Reduce(max, 
                  lapply(X = unnest(cut(as.dendrogram(hc), h=h_i)$upper), 
                         FUN = attr, which = "members") ) # from lukeA's comment

  print(paste0("biggest: ", biggest, ", smallest: ", smallest, " i: ", i, " h_i: ", h_i))
}
h_i

cut(as.dendrogram(hc), h=h_i)
```

How many items in each cluster

```{r}
# histogram
freqs_per_cluster <- 
cuts %>%
  data_frame(cuts = .,
             freq = names(.)) %>% 
  group_by(cuts) %>% 
  tally(sort=TRUE)

hist(freqs_per_cluster$n)

median(freqs_per_cluster$n) # 10
```

```{r}
# put topics back onto abstract spreadsheet
# abstract_ids <- dimnames(general_2018_reduced_dtm)$Docs
# general_2018$`Abstract Id`

general_2018_posters_with_topics <- 
  bind_cols(general_2018_posters, data_frame(cuts))

write.csv(general_2018_posters_with_topics, "../data/derived_data/general_2018_posters_with_clusters.csv")
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```