### 2017-10-27 analysis/paper/saa2018_nonposter_sessions_knapsack.Rmd
while loop seems to be working, we can assign sessions while checking for day-part conflicts

---
title: "Notes on scheduling the SAA2019 meeting"
author:
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  tint::tintHtml: 
     self_contained: TRUE
     df_print: paged
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE, warning = FALSE, message = FALSE,}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(readxl)
library(tidyverse)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")

my_kable <- function(dt){
kable(dt) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE,
                position = "left") %>% 
    column_spec(seq(1, ncol(dt), 2),  background = "grey80")
}
```
## Introduction

In these notes I attempt to make a start on organising the conference sessions into time slots suitable for scheduling.

We have the individual submissions already organised into sessions. The task here is to identify the slots available and fit the sessions into them. Here's what I've done:

First, we computed a suitable duration for each session, given the number of participants and the SAA policy about the maximum duration of the various session types.

Second, we assigned all non-poster sessions to the available four hour time slots. We tallied up the number of slots by looking at the 'Room & Time' spreadsheet.

Third, we spread the slots across the parts of each day of the meeting (morning, afternoon, evening), since each day has slightly different room resources available.

Fourth, we identified problems and schedule clashes, and what to do next.

## How many sessions do we have?

```{r}
library(readxl)
org_sessions <- read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx")

gen_sessions <- read_excel("../data/raw_data/01 - gen_paper_sessions_FINAL(oct5)_DG_including_moves v2.xlsx")

poster_sessions <- 
  org_sessions %>% 
  filter(`Session Type` == "Poster Symposium")

# organised sessions without posters
org_sessions <- 
   org_sessions %>% 
  filter(`Session Type` != "Poster Symposium")

# how many unique organised sessions? 
n_org_sessions <- length(unique(org_sessions$`Session Id`))

# how many unique general sessions? 
n_gen_sessions <- length(unique(gen_sessions$SESSION))

# how many unique poster sessions? 
n_pos_sessions <- length(unique(poster_sessions$`Session Id`))

# total number of sessions
total_number_of_sessions <-  n_org_sessions + n_gen_sessions + n_pos_sessions
```

We have this many sessions in each major category:

```{r}
# table
tribble(       ~'Session type',             ~number,
                      'Organised sessions',        n_org_sessions,
                      'Poster sessions',           n_pos_sessions,
                      'General sessions',          n_gen_sessions)
```

With a total of `r total_number_of_sessions` sessions altogether, and a total number of non-poster sessions `r total_number_of_sessions - n_pos_sessions`.

For scheduling into rooms and time slots, we wont include the posters, because they go in their own schedule in the ballroom.

## How many 4 hour slots do we have available?

To find out how many 4 h slots we have available, I looked on the grid on 'Rooom&Time_DG.xlsx' and counted up the 4  slots. I did not count poster sessions here, because they are less than 4 h and they are outside of the room schedule. I rounded down a little:

- Thur afternoon has the ethics bowl from 1-3 pm, so I removed one whole 4 h slot from that part   
- Fri morning has some time reserved from 9am-12 pm with the note "Reserved for Business Meeting after 12:00pm", so removed one whole 4 h slot from that part also

Here's what I got for session slots in each day and part of day:

```{r}
library(tidyverse)
four_h_slots <- 
  tribble(~part,       ~Th, ~Fr, ~Sa, ~Su,
          "morning",    36,  31,  30,  28,
          "afternoon",  35,  32,  30,  0,
          "evening",    36,  0,    0,  0)

(four_h_slots)
```

```{r}
four_h_slots_long <- 
  four_h_slots %>% 
  gather(times, slot, -part)

# how many 4 h slots
n_four_h_slots <-  sum(four_h_slots_long$slot)
```

So that's a total of `r n_four_h_slots` 4 hour slots, in which we have to fit all the organised and general sessions. This does not include the posters.

## What are the durations of the sessions?

We need to set the upper bounds for the session durations, according to the SAA rules, and then compute how much time each session is likely to take give the number of papers/speakers, if that is less than the allowed time.

The plot below shows the distribution of papers across all non-poster session:

```{r}
# combine organised and general sessions
org_sessions$`Session Id` <- as.character(org_sessions$`Session Id`)
gen_sessions$`Session Id` <- gen_sessions$SESSION
gen_sessions$`Session Type` <- "Symposium"

# get only columns that are also in the organised sessions data frame
gen_sessions_to_join <- 
gen_sessions %>% 
  select(names(gen_sessions)[names(gen_sessions) %in% names(org_sessions)])

all_nonposter_sessions <- 
  bind_rows(gen_sessions_to_join, org_sessions)

# how many papers per session?
papers_per_session <- 
all_nonposter_sessions %>% 
  group_by(`Session Id`) %>% 
  tally(sort = TRUE)

ggplot(papers_per_session,
       aes(n)) +
  geom_histogram() +
  ylab("Papers per \nnon-poster session") +
  theme_bw()
```

Note that I treat chairs and discussants the same as presenters, even though only presenters are really supposed to be allocated program time. The reason for keeping them in is to keep the names so we can avoid schedule clashes where one person is scheduled to be in two rooms at once.

Here are the SAA rules for the duration of each session type:

```{r}

# what's the duration of each session, and he max allowed duration? 
session_types <- unique(all_nonposter_sessions$`Session Type`)

# this comes from 'Conflicts.docx'
session_time_limits <- 
  tribble(~`Session Type`,             ~saa_allowed_time,
          "Poster Symposium",           2,
          "Symposium",                  4,
          "Electronic Symposium",       2, 
          "Forum" ,                     2, 
          "Lightning Rounds",           2)

(session_time_limits)
```

Here we take the SAA rules for the duration of each session type, and if a session has more than 16 papers, we simply assign the duration of that session the maximum allowed time according to the SAA rules. We also compute the 'anticipated duration' of each session by taking the number of papers and mulitplying by 15 minutes. This is most relevant for the 'symposium' types, because other session types don't always have paper presentations. But we apply the upper duration to all sessions for now to ensure they all follow the SAA rules about duration.

```{r}
# join time limits onto session data
session_with_limits_int <- 
  all_nonposter_sessions %>% 
  filter(`Session Type` != "Poster Symposium") %>% 
  left_join(session_time_limits)

# now let's see which session exceed the time limits
min_duration <-  1.5 
round_up_by <- 0.25
session_with_limits <- 
session_with_limits_int %>% 
  group_by(`Session Id`, 
           `Session Type`, 
           saa_allowed_time ) %>% 
  tally(sort = TRUE) %>% 
  mutate(session_time_ideal = n * 15/60) %>% 
  # round up to nearest 0.5 hour
  mutate(session_time_allowed = ceiling(session_time_ideal / round_up_by) * round_up_by) %>% 
  # set the SAA-allowed time as the max time
  mutate(session_time_allowed = if_else(session_time_allowed >= saa_allowed_time,
                                  saa_allowed_time, 
                                  session_time_allowed)) %>% 
  # set a min duration of 2 h
    mutate(session_time_allowed = if_else(session_time_allowed <= min_duration,
                                  min_duration, 
                                  session_time_allowed)) 
```

The plot below shows how a bunch of sessions have had their durations change according the following rules:

- round up each session duration to the nearest `r round_up_by` hour, for convienence, since it's awkward to have sessions starting and ending at odd times    
- shorten sessions to fit into the SAA rules of 2 and 4 h, depending on the session type    
- for very small sessions, set a minumum duration of `r min_duration`, because that's what most session organisers wil be expecting

```{r fig.fullwidth = TRUE, fig.width = 10}

# plot the effect of setting the SAA limits
session_with_limits %>% 
  ungroup() %>% 
  select(`Session Id`, 
         `Session Type`,
         session_time_ideal, 
         session_time_allowed) %>% 
  gather(time_type, time,  
         -`Session Id`, 
         -`Session Type`) %>% 
ggplot(aes(time)) +
  geom_histogram() +
  geom_vline(xintercept = c(2,4), 
             colour = "red") +
  facet_wrap(~   time_type + `Session Type`, ncol = 4) +
  theme_bw() +
  xlab("Duration in hours") +
  ylab("Number of non-poster sessions") + 
  ggtitle("Distribution of durations of non-poster sessions",
          subtitle = paste0("The 'ideal' session times are computed by 15 min x N participants in the session. The 'allowed' times are mandated by the SAA, \nwith ideal times rounded up to the nearest ", round_up_by, " h, and a minimum duration of ", min_duration, "  h "))

```

QUESTION! If there are a very small number of paricipants in a session, do we allocate the SAA-imposed limit to that session? Or do we allow smaller sessions to have shorter durations, for example schedule 1 hour for 4 people? That seems unfair, so I've set a minimum duration of `r min_duration` h for all sessions, since that's what most session chairs will be expecting. I've also rounded up the durations to the nearest `r round_up_by` min

## How to distribute the sessions in the 4 hour slots?

Now we have computed durations for all the sessions, we can try to find the optimum way to squeeze them into the slots available.

This question is combinatorial optimization question, more specifically, it is type of multiple knapsack problem. Given a set of resources (i.e. sessions of various lengths), how do we allocate them to a give number of time slots (all the same size, four hours)? Once we solve that problem, then we will distribute these slots across the day-parts.

To get started on this, we can use Rasmus Bååth's nice case study of the the multiple knapsack problem (<http://www.sumsar.net/blog/2016/06/how-to-cut-your-planks-with-r/>).

```{r}
# Convert the durations to minutes so we get integers, mknapsack doesn't work with decimals

# into these
sessions_we_have <- rep(4, n_four_h_slots) * 60 # so many 4 hour slots that we need to fill 
# we want to fit these
sessions_we_want <- session_with_limits$session_time_allowed * 60 # the duration of each session

# add a buffer onto the sessions less than the full slot
buffer <- 15
sessions_we_want <- ifelse(sessions_we_want <= max(sessions_we_want) - buffer, 
                           sessions_we_want + buffer,
                           sessions_we_want)

library(adagio)
# mknapsack calling signature is: mknapsack(values, weights, capacities)
solution <- mknapsack(sessions_we_want, sessions_we_want, sessions_we_have)
```

Let's put the result back onto the data frame of all the non-poster sessions so we can see which sessions can go in the same slot.

In the snippet below we see that our multiple knapsack function has allocated two sessions into one computed_slot. This looks good because Electronic Symposia and Lightning Rounds can only be 2 hours duration, so we should be able to fit two into one slot. So the multiple knapsack approach seems to be working.

```{r}
# put the computed slot back onto the full details of the sessions
session_with_limits$computed_slot <- solution$ksack

session_with_limits %>% 
  arrange(session_time_allowed, computed_slot) %>% 
  head 
```

Now that we have assigned sessions to slots, we need to join this slot information onto the data frame of each individual paper. This is necessary so we can ensure that we don't put one person in two places at once when we distribute the slots across the day-parts on the schedule.

```{r}
all_nonposter_sessions_with_slots <- 
all_nonposter_sessions %>% 
  left_join(session_with_limits)
```

# How to avoid putting one person in two places at once?

We make a unique ID out of the first name and last name to track individuals.

```{r}
all_nonposter_sessions_with_slots$uid <- 
  with(all_nonposter_sessions_with_slots, paste0(`First Name`, 
                                                 "_", 
                                                 `Last Name`))
```

We assign slots to day-parts, then we join this day-part data back to the data frame of individual papers.

The table below shows a snippet of how the slots available are distributed across the day-parts:

```{r}
four_h_slots_long$cumsum <- cumsum(four_h_slots_long$slot)

xx <- 
four_h_slots_long %>% 
  mutate(day_part = paste0(times, "-", part)) %>% 
  select(-part, -times) %>% 
  rename(end = cumsum) %>% 
  filter(slot != 0)

# get the number of the start slot to go with the number of the end slot
for(i in 1:(nrow(xx))) {
  xx$start[i] <- 
    if_else(i == 1, 
           1,
            as.numeric(xx[ i-1, 2]) + 1)
}

intervals <- 
xx %>% 
  select(-slot) %>% 
  select(day_part, start, end ) %>% 
  rename(from_slot = start,
         to_slot = end)

(head(intervals))
```

Now we take the computed slot values for each session that we got from the multiple knapsack function and put the sessions into day-parts. This table shows a snippet of the computed slots assigned to day-parts:

```{r}
elements <- unique(session_with_limits$computed_slot)

library(fuzzyjoin)
join_intervals_and_elements <- 
fuzzy_left_join(data.frame(elements), 
                intervals, 
                by = c("elements" = "from_slot", 
                       "elements" = "to_slot"), 
                match_fun = list(`>=`, `<=`)) %>% 
  distinct()

computed_slots_in_day_parts <- 
join_intervals_and_elements %>% 
  arrange(elements, day_part) %>% 
  rename(computed_slot = elements)

(head(computed_slots_in_day_parts))
```

Now we join the day-parts onto the individual sessions and then the individual abstracts by the computed slot values.

And now we can plot one possible configuration of the sessions as a grid of sessions for each block of time on the schedule:

```{r fig.fullwidth = TRUE, fig.width = 10, fig.height = 8}
# sessions
all_nonposter_sessions_with_day_parts <- 
session_with_limits %>% 
  left_join(computed_slots_in_day_parts)

# find out what sessions are fractions of 4 hour slots
zzz <- 
all_nonposter_sessions_with_day_parts %>% 
  group_by(computed_slot ) %>% 
  add_count() %>% 
  ungroup() %>% 
  mutate(fraction_of_slot = 1/nn) %>% 
  arrange(computed_slot)

zzzz <- 
zzz %>% 
  group_by(day_part) %>% 
  mutate(day_slot =  computed_slot - from_slot + 1) %>% 
  ungroup()

# get session label string
session_label_string <- 
zzzz %>% 
  group_by(computed_slot) %>% 
  summarise(strings = toString(unique(`Session Id`)))

z6 <- 
  zzzz %>% 
  left_join(session_label_string)

ggplot(z6, 
       aes(day_part, 
           day_slot)) +
  geom_tile(colour = "black", 
           aes(fill = factor(session_time_allowed))) +
  geom_text(aes(label = strings), 
                  size = 3) +
  scale_fill_viridis_d() +
  scale_x_discrete(limits = xx$day_part) +
  scale_y_continuous(breaks = 0:max(z6$day_slot)) +
  guides(colour = FALSE) +
  theme_bw() +
  xlab("") +
  ylab("Room-ish") +
  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
          subtitle = "The numbers on the plot are the Session Ids. \n We can think of each row as equivalent to a room at the conference venue. \n The colour of the cell indicates the duration of the individual session(s) that each cell represents.\n For example the purple cells with two session IDs means that each session is 2 h, for a total of 4 hours in that slot.")

# can we output that as a spreadsheet?
library(tidyr)
z6_plot_as_spreadsheet <- 
  z6 %>% 
  select(day_part, day_slot, strings) %>% 
  distinct(strings, .keep_all = TRUE) %>% 
  spread(day_part, strings, fill = "") %>% 
  select(one_of(xx$day_part))

write_csv(z6_plot_as_spreadsheet,
          "../data/derived_data/visual_schedule_plot_as_spreadsheet.csv")

# papers
all_nonposter_papers_with_day_parts <- 
all_nonposter_sessions_with_slots %>% 
  left_join(computed_slots_in_day_parts)

```

Now let's identify problems where one person is in two different sessions in the same day-part (i.e. column in the plot above).

In the data frame of individual papers, we check that the UID does not occur in more than one session ID for one day-part (it can occur twice on one session ID).

This plot shows sessions where one person is currently scheduled to be in two places at one (the red cells indicate the clashing sessions):

```{r fig.fullwidth = TRUE, fig.width = 10, fig.height = 8}
# identify people with roles in more than one session, but only in the same day-part
# we don't care about mulitple roles in one single session
three_or_more <- 
all_nonposter_papers_with_day_parts %>% 
  group_by(day_part, uid) %>% 
  add_count(sort = TRUE)  %>% 
  filter(nn >= 3) %>% 
  ungroup()  %>% 
  distinct(`Session Id`, uid, .keep_all = TRUE) %>% 
  group_by(uid) %>% 
  filter(n()>1)

# join clashing sessions with plotting data
session_with_clash_flags <- 
  z6 %>% 
  left_join(three_or_more, c("Session Id", "Session Type", "saa_allowed_time", "session_time_ideal", "session_time_allowed", "computed_slot", "day_part", "from_slot", "to_slot")) %>% 
  mutate(fill = if_else(is.na(nn.y), "no_clash", "clash"))

# plot again 
ggplot(session_with_clash_flags, 
       aes(day_part, 
           day_slot )) +
  geom_tile(aes(fill = factor(fill)),
            colour = "black") +
  geom_text(aes(label = strings), 
                  size = 3) +
  scale_x_discrete(limits = xx$day_part) +
  scale_y_continuous(breaks = 0:max(z6$day_slot)) +
  scale_fill_manual(values = c("red", "green")) +
  guides(fill=FALSE, 
         colour = FALSE) +
  theme_bw() +
  xlab("") +
  ylab("Room-ish") +
  ggtitle("Visual schedule for SAA2019 conference non-poster sessions",
          subtitle = "The numbers on the plot are the Session Ids. \n We can think of  each row as equivalent to a room at the conference venue. \n Red cells indicate that one person is currently scheduled to be in both sessions :(")

```

## How can we prevent one person scheduled in two simultaneous sessions?

If we treat each day-part as one knapsack, and then loop over filling the knapsack to check if a name has already been added, and if it has, skip that day-part and move to the next, and check that next pay-part to see if the name is already there.

We can compute the total capacity of each day-part in hours like this:

```{r}
# each slot is 4 hours long
total_cap <- 
  xx %>% 
  mutate(total_capacity = slot * 4) %>% 
  select(day_part, slot, total_capacity)
total_cap
```

Now we will take session from the big table of sessions and add them into, for example, Th-morning until we reach the capacity of that day-part. But, we also need to loop over the sessions as we add more to the day-part to check to see if a person's name is already there or not.

```{r echo = TRUE, eval = FALSE}

# get all the sessions, but we're going to modify this, so make a copy
z7 <- z6
all_of_the_day_parts <- NULL
total_cap$hours_of_sessions <- 0
total_cap$n_sessions <- 0
total_cap$c_sessions <- 0
container_for_this_day_part_with_day_from_while <- 0
i <- 1

# start with each day-part...
for(i in seq_len(nrow(total_cap))) {
  target_capacity <- total_cap$total_capacity[i] - 4

  # prepare a while statement that will take a session, 
  # check to see if there is a name clash with sessions,
  # skip adding it if there is clash, then check to see
  # the culumative duration of the sessions in the day-part 
  # has exceeeded the total duration of the day-part,
  # and stop adding sessions if it has.

# make sure some variables are empty
cumulative_duration_of_sessions   <-  0
container_for_this_day_part <- 0
all_the_people_in_the_session <- 0
j <- 1

while(cumulative_duration_of_sessions <= target_capacity & 
      j <= nrow(z7)) {

    get_a_session <- z7[j, ]

    # get the names of all the people in this session...
    all_the_people_in_this_session <- 
      left_join(get_a_session, 
                all_nonposter_papers_with_day_parts, 
                by = "Session Id")

    # get names from what we already have added to this day-part
     if (j == 1) {
      # do nothing is this is the first incoming session 
       distinct_names_for_each_session_so_far <- 0
    } else {
      # get distinct names for each session previously added
      # to this day-part
      # (don't care about same name multiple times in one session)
      distinct_names_for_each_session_so_far <- 
      container_for_this_day_part %>% 
        group_by(uid) %>% 
        distinct(uid, .keep_all = TRUE) %>% 
        pull(uid)
    }

    # compare names already in here to names we are about to add
    if (j == 1) {
      # do nothing 
      names_already_in <- 0
      } else {
      # check to see if incoming session has names that already exist in this day_part
      names_already_in <- sum(all_the_people_in_this_session$uid %in% distinct_names_for_each_session_so_far)
      }

    # accumulate the incoming session onto the sessions
    # we've already added to this day-part, if there are no
    # name conflicts
    if (j == 1) {
       container_for_this_day_part <- 
        all_the_people_in_this_session
      } else {
      if (names_already_in == 0) {
      # if there are no name clashes, 
      # add these names to previous names in this day-part

      # bind rows of session that has no clashing  names with 
      # sessions we already added to this day-part
      container_for_this_day_part <- 
        bind_rows(all_the_people_in_this_session,
                  container_for_this_day_part)
    }  else {

        # do not add this session to the day-part
      }
    }

    # compute the duration of session that we've added so far to 
    # this day-part
    cumulative_duration_of_sessions <- 
      container_for_this_day_part %>% 
      distinct(`Session Id`, .keep_all = TRUE) %>% 
      pull(session_time_allowed.x) %>% 
      sum

    # store this result
    total_cap$hours_of_sessions[i] <- cumulative_duration_of_sessions

    # show a check of what's happening
    if (j == 1){
    progress_check <-    
      data_frame(i = i, 
                 j = j,
               'cumulative duration' = cumulative_duration_of_sessions,
               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
               people =  nrow(container_for_this_day_part)
                )
    } else {
      progress_check <- bind_rows(progress_check, 
                                  data_frame(i = i,
                                             j = j,
               'cumulative duration' = cumulative_duration_of_sessions,
               'sessions added' =  length(unique(container_for_this_day_part$`Session Id`)),
               people =  nrow(container_for_this_day_part)
                ))
    }

       # print(progress_check)

       # move to next session
       j <- j + 1
    } 
# end of the while loop

 # add an identifier so we can see where we have assigned each session so far
 container_for_this_day_part_with_day_from_while <- 
   container_for_this_day_part %>% 
   mutate(day_part_from_while  = as.character(total_cap$day_part[i]))

 # bind rows to previous day-parts we might have done
  if (i == 1) {

  all_of_the_day_parts <- container_for_this_day_part_with_day_from_while

  } else {

    # check to see

  all_of_the_day_parts <- bind_rows(container_for_this_day_part_with_day_from_while, 
                                      all_of_the_day_parts)
  }

 # progress check
 print(paste0("i = ", i))

 print(
 all_of_the_day_parts %>% 
   group_by(day_part_from_while) %>% 
   summarise(papers = n(),
             sessions = length(unique(`Session Id`)))
 )

 total_cap$n_sessions[i] <- length(unique(container_for_this_day_part_with_day_from_while$`Session Id`))

 total_cap$c_sessions[i] <-  sum(total_cap$n_sessions)

print(total_cap)

 # for the next day-part, we need to make sure
 # we don't add the same session twice to different day-parts
 # so let's update our big table of abstracts, and drop the session that 
 # we've already got in the day-part

 z7 <-  
   z7 %>% 
   filter(! `Session Id` %in% unique(container_for_this_day_part$`Session Id`))

 if (nrow(z7) == 0) stop("Stopping the loop because all sessions have been assigned to day-parts")

}
```

```{r}
# check to see what we got from the loop...

n_missing_papers <- 
  nrow(all_nonposter_sessions_with_slots) - nrow(all_of_the_day_parts)

n_missing_sessions <- 
  length(unique(all_nonposter_sessions_with_slots$`Session Id`)) - length(unique(all_of_the_day_parts$`Session Id`))

missing_sessions <- 
  unique(all_nonposter_sessions_with_slots$`Session Id`)[!(unique(all_nonposter_sessions_with_slots$`Session Id`) %in% unique(all_of_the_day_parts$`Session Id`)) ]

# have we succeeded with no name clashes?
# in each day-part, do any names occur more than once?
clash <- 
all_of_the_day_parts %>% 
  group_by(`Session Id`, day_part_from_while, uid) %>% 
  tally() %>% 
  arrange(`Session Id`, n)

```

## Summary

Now we have a rough draft of the schedule of non-poster sessions. We used the multiple knapsack algorithm to distribute the sessions efficiently across the available sessions. This method identifies combinations of sessions that are <4 h and then combines these shorter sessions to fill the 4 h slots.

We still have some things to do:

- Resolve the clashes that currently have one person in two sessions at the same time. I'm thinking of how it can be done programatically, but nothing obvious has occured to me.     
- Assign the sessions to specific times and rooms. Perhaps this is best be done manually?    
- Consider all of the conflicts and special requests in 'Conflicts.docx'    
- Repeat all of this for the posters. This could be done with some minor variations to the code used here.

<!--

## Working in a local Docker container

To run this project in a local Docker container, I start a bash shell in the project directory:

Do this one time only, build my Docker container from Dockerfile at top-level in my project:

```
docker build - < Dockerfile
docker ps
```

Take note of the container ID, which you get after `docker ps`

Then run my Docker container, access it via the web browser, link the project dir to the Docker drive. Make sure you're in a local directory that is sharable (C:\Users\yourname is usually good):

```
docker run -dp 8787:8787 -v ${pwd}:/home/rstudio/ -e ROOT=TRUE  <container ID>
```

Then go to `http://192.168.99.100:8787/` or `localhost:8787` in youor browser, log in (rstudio/rstudio), and start your RStudio project by double-clickin on the .Rproj file

When you're done, in the shell, stop all docker containers with:

```
docker stop $(docker ps -a -q)
```

Or do `docker ps` and `docker stop <container ID>` just to stop one container. 
-->

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```