### 2017-11-03 analysis/paper/saa2018_posters.Rmd
making a start on the posters

---
title: "Notes on scheduling the SAA2019 meeting"
author:
  - Ben Marwick
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  tint::tintHtml: 
     self_contained: TRUE
     df_print: paged
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r, setup, echo = FALSE, warning = FALSE, message = FALSE,}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "figures"
)

library(saaabstracts)
library(readxl)
library(tidyverse)
library(knitr)
library(glue)

```
## Introduction

## How many sessions do we have?

```{r}
library(readxl)

gen_poster_session_posters <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx")

org_poster_sessions_only <- read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", sheet = "ORGANIZED_POSTER_SESSION_SIZES")

after_hours_poster_sessions_only <- 
  read_excel("../data/raw_data/02 - poster_sessions_FINAL(oct31).xlsx", 
             sheet = "AFTER_HOURS") %>% 
  slice(1:9)

# In here we have org and after hours posters
org_and_after_hours_posters_individual <- 
  read_excel("../data/raw_data/Organized Session Abstracts_2018_AB.xlsx") %>% 
  filter(`Session Type` == "Poster Symposium")

# how many unique general sessions? 
n_gen_poster_sessions <- length(unique(gen_poster_sessions$cuts))

# how many unique organised sessions? 
n_org_poster_sessions <- nrow((org_poster_sessions))

# how many unique after hours sessions? 
n_ah_poster_sessions <- nrow(after_hours_poster_sessions_only)

# total number of sessions
total_number_of_sessions <-  n_gen_poster_sessions + n_org_poster_sessions + n_ah_poster_sessions

# combine organised and general sessions
gen_poster_session_posters$session <- gen_poster_session_posters$cuts

org_poster_session_posters <- 
  org_and_after_hours_posters_individual %>% 
  mutate(session = `Session Id`) %>% 
  filter(session %in% org_poster_sessions_only$ID)

common_cols <- intersect(names(gen_poster_session_posters),  names(org_poster_session_posters))

gen_poster_session_posters <- 
  gen_poster_session_posters %>% 
  select(common_cols)

org_poster_session_posters <- 
  org_poster_session_posters %>% 
  select(common_cols)

gen_and_org_poster_session_posters <- 
  bind_rows(gen_poster_session_posters, 
            org_poster_session_posters)

```

We have information about how many posters can go into each session

```{r}
# poster session durations
poster_session_durations <- 
  read_excel("../data/raw_data/saa all organizedsession_timeslotABv3.xls", 
             sheet = "Posters1")
```

What do we stil need to do?

- Posters

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
git2r::repository("../..")
```